---
title: "Class12: Structural Bioinformatics pt.1"
output: github_document
---

##PDB statistics

Here we inspect the type of structures in the main database for 3D biomolecular data - the PDB

>Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy.

```{r}
#read in file
stats <- read.csv("Data Export Summary.csv", row.names = 1)

#calculate percentage values asked for each "method"
ans <- stats$Total/ sum(stats$Total) *100
names(ans) <- rownames(stats)
round(ans, 2)

```


>Q2: Also can you determine what proportion of structures are protein? Aim to have a rendered GitHub
document with working code that yields your answers.

```{r}
protein <- stats$Proteins/ sum(stats$Proteins) *100
names(protein) <- rownames(stats)
round(protein, 2)
```


>Q3: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

