---
title: "Class18: Investigating cancer genomics datasets"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Connect to the NCI-GDC database

The NCI-GDC (cancer genomic commons) is a major protal for cancer genomic data.

Here we use the bioconductor package **GenomicDataCommons** to talk to this database and download data etc. all from R.
```{r}
library(GenomicDataCommons)
library(TCGAbiolinks)
library(maftools)
```

Can we talk to the online database?
```{r}
status()
```

##Querying the GDC from R
```{r}
cases_by_project <- cases() %>%
  facet("project.project_id") %>%
  aggregations()
head(cases_by_project)
```

*The pipe operator can be useful in R and sometimes is more clear to read.*
**c(1:10) %>% sum() == sum(c(1:10))**


###Plot this overview data
```{r}
x <- cases_by_project$project.project_id
barplot(x$doc_count, names.arg = x$key, las=2, col="lightblue", log="y")
```



##Variant analysis with R

Find files available for PAAD project that we can then examine variants from
```{r}
maf.file <- GDCquery_Maf(tumor="PAAD", pipelines = "mutect")
head(maf.file)
```

Read in the MAF file we found
```{r}
vars = read.maf(maf = maf.file, verbose = FALSE)
plotmafSummary(vars)
```

##Oncoplots
```{r}
oncoplot(maf = vars, top = 10)
```

