---
title: "Class17"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Setup
First setup the packages we need. These include **igraph** from CRAN and **RCy3** from bioconductor. **install.packages("igraph")** and **BiocManager::install("RCy3")**

```{r}
library(RCy3)
library(igraph)
```

Let's check we can talk to cytoscape from R. Note that cytoscape must be running at this point.
```{r}
cytoscapePing()
```


We can test things further by making a small network (here in igraph format as used by the R igraph package) and sending it to Cytoscape:
```{r}
g <- makeSimpleIgraph()
createNetworkFromIgraph(g,"myGraph")
```

```{r}
setVisualStyle("Universe")
```


```{r}
## scripts for processing located in "inst/data-raw/"
prok_vir_cor <- read.delim("virus_prok_cor_abundant.tsv", stringsAsFactors = FALSE)

## Have a peak at the first 6 rows
head(prok_vir_cor)

```


How many connections (i.e. edges) will we have in our network - lets look at the number of rows
```{r}
nrow(prok_vir_cor)
```

```{r}
g <- graph.data.frame(prok_vir_cor, directed = FALSE)
```

```{r}
class(g)
```

```{r}
plot(g)
plot(g, vertex.size=3, vertex.label=NA)
```


##Network community detection

Lets cluster our network to find

Community structure detection algorithms try to find dense sub-graphs within larger network graphs (i.e. clusters of well connected nodes that are densely connected themselves but sparsely connected to other nodes outside the cluster) . Here we use the classic Girvan & Newman betweenness clustering method. The igraph package has lots of different community detection algorithms (i.e. different methods for finding communities).


```{r}
cb <- cluster_edge_betweenness(g)

plot(cb, y=g, vertex.label=NA,  vertex.size=3)
```

```{r}
pr <- page_rank(g)
head(pr$vector)

v.size <- BBmisc::normalize(pr$vector, range=c(2,20), method="range")
plot(g, vertex.size=v.size, vertex.label=NA)
```

