---
title: "Class10"
author: "Nicholas Chen"
date: "2/6/2020"
output: html_document
---

#Analysis of Human Breast Cancer Cells

First we need to import (i.e. read) our input data. I downloaded the CSV file from the class website: https://bioboot.github.io/bimm143_S18/class-material/WisconsinCancer.csv

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

There are some funkey things in this dataset that we will ignore for our analysis. This includes the first and second ID and Diagnosis columns and the funny last X column (col 33).

```{r}
wisc.data <- as.matrix( wisc.df[,3:32] )
head(wisc.data)
```

Q. How many patients do we have data for?

```{r}
nrow(wisc.data)
```

Q. How many cancer and non-cancer?

```{r}
table(wisc.df$diagnosis)
```

Q. How many variables/features in the data are suffixed with _mean?

For this I will turn to the `grep()` function and look at the help page to see how it works.

```{r}
colnames(wisc.data)
#to look at the column name
```

```{r}
#grep(pattern, vector)
grep("_mean", colnames(wisc.data), value = TRUE)
```

```{r}
grep("_mean", colnames(wisc.data))
# number of outcome columns
```

```{r}
length( grep("_mean", colnames(wisc.data) ) )
# number of total outcome columns
```

## Principal Component Analysis

Before we turn to PCA we need to think, or consider, whether we should SCALE our input.

It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

The input variables use different units of measurement.

- The input variables have significantly different variances.
- Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the `colMeans()` and `apply()` functions like you’ve done before.

```{r}
round( apply(wisc.df, 2, sd), 2 )
#round to 2 sig.fig.
```

Look like we need to set scale=TRUE

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp( wisc.data, scale=TRUE )
summary(wisc.pr)

# Scale: if TRUE, the centered column values are divided by the column’s standard deviation (when center is also TRUE; otherwise, the root mean square is used); if FALSE, scaling is not performed
```

Q. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

The 1st PC captures 44.27% of the original variance. Note that 72.6% is captured in the first 3 PCs..

Lets make some figures...

```{r}
biplot(wisc.pr)
```

That is a hot mess! We need to do our own PC1 vs PC2 plot and lets color by the diagnosis.

```{r}
attributes(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
abline(h=0, col="gray", lty=2)
abline(v=0, col="gray", lty=2)
```

###CLuster in PC space

cluster the original data

#use
```{r}
kmeans(wisc.data, centers = 2, nstart = 20)
```

#or
```{r}
wisc.hc <- hclust( dist(wisc.data) )
plot(wisc.hc)
```

This does not look good! Let's try and combine the results of PCA with clustering...

Using the minimum number of principal components required to describe at least 90% of the variability in the data, create a hierarchical clustering model with the linkage method="ward.D2". We use Ward’s criterion here because it is based on multidimensional variance like principal components analysis. Assign the results to wisc.pr.hclust.

```{r}
wisc.pr.hclust <- hclust( dist(wisc.pr$x[,1:3]), method="ward.D2" )
plot(wisc.pr.hclust)
```

To get our clusters out of this tree we need to CUT it with the `cutree()` function.

```{r}
grps3 <- cutree(wisc.pr.hclust, k=3)
table(grps3)

```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps3)

```

We can use the `table()` function to compare the $diagnosis vector with our cluster results vector.

```{r}
table(grps3, wisc.df$diagnosis)
```

##Prediction

```{r}
new <- read.csv("new_samples.csv")
new
```

Use `predict()` function with our previous PCA model and new data...
```{r}
npc <- predict(wisc.pr, newdata=new)
npc
```

Now draw the PCA plot again and add our new data:

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=15, cex=2)
text(npc[,1],npc[,2], labels=c(1,2), col="white")
```

